---
title: "Getting water level and temperature"
author: "Karine Villeneuve"
date: "07/12/2022"
output: html_document
---
Data for groundwater level and temperature was collected and provided to me by the laboratory of Marie Larocque as well as Florent Barbecot at the University of Quebec in Montreal.
The files contained all the groundwater temperature and level above the probe for each hour of the day from January 1th 2019 until December 31 2020. I used this script in order to extract only the data from the day I sampled and to remove outliers produced by sampling. Finally, the hourly measured were averaged to obtain only one measure per day (daily average)
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/karinevilleneuve/OneDrive - UQAM/DOCTORAT/Projects/Chapitre-1_Spatial/Hydrogeo_data/")
library(ggplot2)
library(dplyr)
library(ggpubr)
``` 
Load data sets
```{r}
piezo_r=read.table("piezo_r.csv", sep=",",header=TRUE) # measures from the Rigaud probe
piezo_sl=read.table("piezo_sl.csv", sep=",",header=TRUE) # measures from the Saint-Lazare probe
```

```{r}
piezo_r$City="Rigaud"
piezo_sl$City="Saint-Lazare"
piezo=rbind(piezo_r,piezo_sl)
piezo$City_date=paste(piezo$City, piezo$Date, sep="_") # Create a new column combining dates and city 
piezo$Date=as.Date(piezo$Date) # Change column to Date format
piezo$month=months(piezo$Date) # Extract only the month
piezo$year=gsub("-.*","",piezo$Date) # Combine month and year
piezo$Month_year=paste(piezo$month, piezo$year, sep=" ")
``` 
Remove daily outliers for groundwater level and temperature 
```{r}
# Remove daily outliers 
temp_list_daily=list()
lev_list_daily=list()
i = 0
for (each_date in unique(piezo$City_date)){
  i=i+1
  date_df=subset(piezo, City_date==each_date)
  # Remove outliers for temperature 
  temp_quartiles=quantile(date_df$TEMPERATURE, probs=c(.25, .75), na.rm=FALSE)
  temp_IQR=IQR(date_df$TEMPERATURE)
  temp_Lower=temp_quartiles[1] - 1.5*temp_IQR
  temp_Upper=temp_quartiles[2]+1.5*temp_IQR
  temp_no_out=subset(date_df, date_df$TEMPERATURE > temp_Lower & date_df$TEMPERATURE < temp_Upper)
  temp_list_daily[[i]]=temp_no_out
  # Remove outliers for level
  lev_quartiles=quantile(date_df$LEVEL, probs=c(.25, .75), na.rm=FALSE)
  lev_IQR=IQR(date_df$LEVEL)
  lev_Lower=lev_quartiles[1] - 1.5*lev_IQR
  lev_Upper=lev_quartiles[2]+1.5*lev_IQR
  lev_no_out=subset(date_df, date_df$LEVEL > lev_Lower & date_df$LEVEL < lev_Upper)
  lev_list_daily[[i]]=lev_no_out
}
temp_daily=do.call("rbind",temp_list_daily)
level_daily=do.call("rbind",lev_list_daily)
```
Remove monthly outliers from groundwater level and temperature 
```{r}
# Temperature 
temp_list_montly=list()
i = 0
for (each_date in unique(temp_daily$Month_year)){
  i=i+1
  date_df=subset(temp_daily, Month_year==each_date)
  temp_quartiles=quantile(date_df$TEMPERATURE, probs=c(.25, .75), na.rm=FALSE)
  temp_IQR=IQR(date_df$TEMPERATURE)
  temp_Lower=temp_quartiles[1] - 1.5*temp_IQR
  temp_Upper=temp_quartiles[2]+1.5*temp_IQR
  temp_no_out=subset(date_df, date_df$TEMPERATURE > temp_Lower & date_df$TEMPERATURE < temp_Upper)
  temp_list_montly[[i]]=temp_no_out
}
temp_df=do.call("rbind",temp_list_montly)
# Level
level_list_montly=list()
i = 0

for (each_date in unique(level_daily$Month_year)){
  i=i+1
  date_df=subset(level_daily, Month_year==each_date)
  lev_quartiles=quantile(date_df$LEVEL, probs=c(.25, .75), na.rm=FALSE)
  lev_IQR=IQR(date_df$LEVEL)
  lev_Lower=lev_quartiles[1] - 1.5*lev_IQR
  lev_Upper=lev_quartiles[2]+1.5*lev_IQR
  lev_no_out=subset(date_df, date_df$LEVEL > lev_Lower & date_df$LEVEL < lev_Upper)
  level_list_montly[[i]]=lev_no_out
}
level_df=do.call("rbind",level_list_montly)
``` 
Combine hourly measures in daily average
```{r}
# Combine but keep columns 
mean_daily_temperature=temp_df %>% 
  group_by(City_date) %>% 
  mutate(mean_daily_temperature=mean(TEMPERATURE)) %>% 
  ungroup() %>%
  distinct(City_date, .keep_all = TRUE) # Remove duplicate


mean_daily_level=level_df %>% 
  group_by(City_date) %>% 
  mutate(mean_daily_level=mean(LEVEL)) %>% 
  ungroup() %>%
  distinct(City_date, .keep_all = TRUE) # Remove duplicate

# Remove columns we don't need anymore
mean_daily_level=subset(mean_daily_level, select=-c(Time, ms, TEMPERATURE, LEVEL))
mean_daily_temperature=subset(mean_daily_temperature, select=-c(Time, ms, LEVEL, TEMPERATURE))
```
Order levels chronologically 
```{r}
mean_daily_temperature$Month_year=factor(mean_daily_temperature$Month_year, levels=c("January 2019","February 2019","March 2019","April 2019",
                                                       "May 2019","June 2019","July 2019","August 2019","September 2019",
                                                       "October 2019","November 2019","December 2019","January 2020",
                                                       "February 2020","March 2020","April 2020","May 2020","June 2020",
                                                       "July 2020","August 2020","September 2020","October 2020",
                                                       "November 2020","December 2020"))
mean_daily_level$Month_year=factor(mean_daily_level$Month_year, levels=c("January 2019","February 2019","March 2019","April 2019",
                                                       "May 2019","June 2019","July 2019","August 2019","September 2019",
                                                       "October 2019","November 2019","December 2019","January 2020",
                                                       "February 2020","March 2020","April 2020","May 2020","June 2020",
                                                       "July 2020","August 2020","September 2020","October 2020",
                                                       "November 2020","December 2020"))
# Seperate dataframe according to city
rigaud_temperature=subset(mean_daily_temperature, City=="Rigaud")
rigaud_level=subset(mean_daily_level, City=="Rigaud")
saintlaz_temperature=subset(mean_daily_temperature, City=="Saint-Lazare")
saintlaz_level=subset(mean_daily_level, City=="Saint-Lazare")
```
Plot the results
```{r}
# plotting results for Saint-Lazare (a)
sl_temp_plot= ggplot(saintlaz_temperature, aes(Month_year, y=mean_daily_temperature)) + 
  geom_boxplot(outlier.color=NA) +
  labs(title="Saint-Lazare", y="Temperature (°C)") +
  theme_classic() +
  scale_y_continuous(limits=c(0,12), expand = c(0,0), n.breaks = 5) + # remove space between 0 and x axis
  theme(text = element_text(size=12, face="plain"), 
        panel.grid.major.x = element_line(color = "gray", size = 0.2,linetype = 3), 
        legend.justification = "center",
        legend.position="bottom", 
        plot.title = element_text(hjust =0.5, size=12, face="plain"),
        axis.text.x = element_blank(), 
        axis.ticks.x=element_blank(), 
        axis.title.x = element_blank()) +
  annotate("rect", xmin="September 2019",xmax = "August 2020",ymin =0, ymax = Inf,
           alpha =0.1,fill = "blue") # Add blue rectangle representing sampling period 

sl_level_plot= ggplot(saintlaz_level, aes(Month_year, y=mean_daily_level)) + 
  geom_boxplot(outlier.color=NA) +
  labs(y="Level (m)") +
  theme_classic() +
  scale_y_continuous(limits=c(0,4), expand = c(0,0), n.breaks = 5) + # remove space between 0 and x axis
  theme(text = element_text(size=12, face="plain"),
        panel.grid.major.x = element_line(color = "gray", size = 0.2,linetype = 3), 
        legend.justification = "center",
        legend.position="bottom", 
        plot.title = element_text(hjust =0.5, size=12,face="plain"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=9), 
        axis.title.x=element_blank()) + 
  annotate("rect", xmin="September 2019",xmax = "August 2020",ymin =0, ymax = Inf,
           alpha =0.1,fill = "blue") # Add blue rectangle representing sampling period 


r_level_plot= ggplot(rigaud_level, aes(Month_year, y=mean_daily_level)) + 
  geom_boxplot(outlier.color=NA) +
  labs(x= "Date") +
  theme_classic() +
  scale_y_continuous(limits=c(0,4), expand = c(0,0)) + # remove space between 0 and x axis
  theme(text = element_text(size=12,face="plain"),
        panel.grid.major.x = element_line(color = "gray", size = 0.2,linetype = 3), 
        legend.justification = "center",
        legend.position="bottom", 
        plot.title = element_text(hjust =0.5, size=12,face="plain"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=9),
        axis.title.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title.x=element_blank()) +
  annotate("rect", xmin="September 2019",xmax = "August 2020",ymin =0, ymax = Inf,
           alpha =0.1,fill = "blue") # Add blue rectangle representing sampling period 

r_temp_plot= ggplot(rigaud_temperature, aes(Month_year, y=mean_daily_temperature)) + 
  geom_boxplot(outlier.color=NA) +
  labs(title="Rigaud") +
  theme_classic() +
  scale_y_continuous(limits=c(0,12), expand = c(0,0), n.breaks = 5) + # remove space between 0 and x axis
  theme(text = element_text(size=12, face="plain"),
        panel.grid.major.x = element_line(color = "gray", size = 0.2,linetype = 3), 
        legend.justification = "center",
        legend.position="bottom", 
        plot.title = element_text(hjust =0.5, size=12, face="plain"),
        axis.text.x = element_blank(), 
        axis.ticks.x=element_blank(), 
        axis.title.x = element_blank(),
        axis.title.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) +
  annotate("rect", xmin="September 2019",xmax = "August 2020",ymin =0, ymax = Inf,
           alpha =0.1,fill = "blue") # Add blue rectangle representing sampling period 

combined_plot=ggarrange(sl_temp_plot, r_temp_plot, sl_level_plot, r_level_plot, ncol=2, nrow=2, labels=c("a","b","c","d"),
                        heights=c(0.8,1), widths=c(0.9,1), align="v") + 
  annotate("text", x=0.521, y=0.012, label = "Month", size = 4)
```

```{r}
ggsave("temperature_level.pdf",plot=combined_plot, device="pdf", path="/Users/karinevilleneuve/OneDrive - UQAM/DOCTORAT/Projects/Chapitre-1_Spatial/images/", width=16.5, height = 16.5, units = "cm")
```

# Generate table with daily average from our sampling period 
```{r}
mdl=subset(mean_daily_level, subset=Month_year %in% c("September 2019","October 2019","November 2019",
                                                   "December 2019","January 2020","February 2020",
                                                   "March 2020","April 2020","May 2020","June 2020",
                                                    "July 2020","August 2020"))
mdt=subset(mean_daily_temperature, subset=Month_year %in% c("September 2019","October 2019","November 2019",
                                                   "December 2019","January 2020","February 2020",
                                                   "March 2020","April 2020","May 2020","June 2020",
                                                    "July 2020","August 2020"))
```

Save tables
```{r}
write.csv(mdt, "/Users/karinevilleneuve/OneDrive - UQAM/DOCTORAT/Projects/Chapitre-1_Spatial/Github/main_daily_temperature.csv", quote=FALSE, row.names = FALSE)
write.csv(mdl, "/Users/karinevilleneuve/OneDrive - UQAM/DOCTORAT/Projects/Chapitre-1_Spatial/Github/main_daily_level.csv", quote=FALSE, row.names = FALSE)
``` 
